---
title: "TTR Amyloidosis (Random Forest Approch)"
output: html_document
---

```{r, echo=TRUE}
library(readr)
library(dplyr)
library(ggplot2)
library(randomForest) 

setwd("~/Dropbox/Gene_Signatures_Code")
```

## Read in data from directory & merge data files

```{r, echo=TRUE}
geneExpressionData <- read_tsv("GSE67784.txt.gz")
colnames(geneExpressionData)[1] <- "SampleID"

metaData <- read_csv("Metadata/Ensembl87_Biomart_GRCh38.p7.csv")
colnames(metaData)[1] <- "Gene"
colnames(metaData)[2] <- "Chromosome"
colnames(metaData)[3] <- "GeneSymbol"

# Remove unnecessary column
metaData <- select(metaData, -4)

metaData <- filter(metaData, metaData$Chromosome %in% c(1:22, "X", "Y"), !is.na(metaData$GeneSymbol))
# filteredMetaData <- filter(metaData, metaData$Chromosome %in% c(1:22))

metaData$Chromosome <- factor(metaData$Chromosome)
# filteredMetaData$Chromosome <- factor(filteredMetaData$Chromosome)
```

## Write a function to create separate datasets

```{r,}
getClassData <- function(geneExpressionData, classFile) {
  classData <- read_tsv(classFile)
  classData$Class <- factor(classData$Class)

  mergedData <- inner_join(classData, geneExpressionData)

  return(mergedData)
}
```

## Age Data
```{r}
ageData <- read_tsv("PatientCharacteristic__age.txt")

```

## Create separate datasets, combine them

```{r}
asymptomaticMergedData <- getClassData(geneExpressionData, "Class/Asymptomatic__gender.txt")
symptomaticMergedData <- getClassData(geneExpressionData, "Class/Symptomatic__gender.txt")
treatedMergedData <- getClassData(geneExpressionData, "Class/Treated__gender.txt")
onlycontrolsMergedData <- getClassData(geneExpressionData, "Class/Only_Controls__gender.txt")

asymptomaticData <- cbind("A", asymptomaticMergedData)
colnames(asymptomaticData)[1] <- "Type"
symptomaticData <- cbind("S", symptomaticMergedData)
colnames(symptomaticData)[1] <- "Type"
treatedData <- cbind("T", treatedMergedData)
colnames(treatedData)[1] <- "Type"
controlsData <- cbind("C", onlycontrolsMergedData)
colnames(controlsData)[1] <- "Type"

combData <- rbind(asymptomaticData, symptomaticData, treatedData, controlsData)
fullData <- inner_join(ageData, combData)

############### I don't have the age data for the treated patients #################
```


## Get p-values for ANOVA
```{r}
nGenes <- ncol(fullData) - 4
geneMatrix <- as.matrix(fullData[,5:nGenes + 4], ncol=nGenes)

p1 <- rep(NA,nGenes)
p2 <- rep(NA,nGenes)
p3 <- rep(NA,nGenes)
p4 <- rep(NA,nGenes)
p5 <- rep(NA,nGenes)
p6 <- rep(NA,nGenes)

x1 <- fullData$age
x2 <- fullData$Class
x3 <- fullData$Type

for (i in 1:ncol(geneMatrix)) {
    y <- geneMatrix[,i] 
    mod <- lm(y~x1+x2+x3+x1*x2+x2*x3+x1*x3)
    p1[i] <- anova(mod)[1,5]
    p2[i] <- anova(mod)[2,5]
    p3[i] <- anova(mod)[3,5]
    p4[i] <- anova(mod)[4,5]
    p5[i] <- anova(mod)[5,5]
    p6[i] <- anova(mod)[6,5]
}

results <- list(p1,p2,p3,p4,p5,p6)

```


## Assign PValues to Genes
```{r, }
geneNames <- colnames(fullData)[5:(ncol(fullData))] 
  
pValueData1 <- data.frame(Gene=geneNames,p1)  # age
pValueData2 <- data.frame(Gene=geneNames,p2)  # class
pValueData3 <- data.frame(Gene=geneNames,p3)  # type
pValueData4 <- data.frame(Gene=geneNames,p4)  # age and class
pValueData5 <- data.frame(Gene=geneNames,p5)  # class and type
pValueData6 <- data.frame(Gene=geneNames,p6)  # age and type

pValueData1$Gene <- as.character(pValueData1$Gene)
pValueData2$Gene <- as.character(pValueData2$Gene)
pValueData3$Gene <- as.character(pValueData3$Gene)
pValueData4$Gene <- as.character(pValueData4$Gene)
pValueData5$Gene <- as.character(pValueData5$Gene)
pValueData6$Gene <- as.character(pValueData6$Gene)

pValueDF1 <- inner_join(metaData, pValueData1)
pValueDF1 <- arrange(pValueDF1, p1)
colnames(pValueDF1)[4] <- "PValue"
pValueDF2 <- inner_join(metaData, pValueData2)
pValueDF2 <- arrange(pValueDF2, p2)
colnames(pValueDF2)[4] <- "PValue"
pValueDF3 <- inner_join(metaData, pValueData3)
pValueDF3 <- arrange(pValueDF3, p3)
colnames(pValueDF3)[4] <- "PValue"
pValueDF4 <- inner_join(metaData, pValueData4)
pValueDF4 <- arrange(pValueDF4, p4)
colnames(pValueDF4)[4] <- "PValue"
pValueDF5 <- inner_join(metaData, pValueData5)
pValueDF5 <- arrange(pValueDF5, p5)
colnames(pValueDF5)[4] <- "PValue"
pValueDF6 <- inner_join(metaData, pValueData6)
pValueDF6 <- arrange(pValueDF6, p6)
colnames(pValueDF6)[4] <- "PValue"

```


## Make Barplot for p1
```{r}
createBarplot <- function(pValueDF) {
  
  tempDataFrame <- subset(pValueDF, select=c("Chromosome", "Gene", "PValue"))

  for(i in 1:nrow(tempDataFrame)) {
      ifelse(tempDataFrame$PValue[i]<=.05, tempDataFrame$PValue[i] <- "Yes", tempDataFrame$PValue[i] <- "No")
  }

  df <- data.frame(tempDataFrame)
  df <- group_by(df, Chromosome, PValue)

  propPlot <- ggplot(df, aes(x=Chromosome, fill=PValue)) + geom_bar() + ggtitle("Significant Genes") 
  return(propPlot)
}

createBarplot(pValueDF1)
createBarplot(pValueDF2)
createBarplot(pValueDF3)
createBarplot(pValueDF4)
createBarplot(pValueDF5)
createBarplot(pValueDF6)
```


## Decision Trees using most sig genes
```{r}
getRate1 <- function() {
  pValueDF1$PValue <- pValueDF1$PValue[order(pValueDF1$PValue)]
  sbst <- pValueDF1$Gene[1:1000]
  fullDataSub <- fullData[,c("age", "Type", "Class", sbst)] 
  fullDataSub$Type <- as.factor(as.character(fullDataSub$Type))

  fitGE <- randomForest(fullDataSub$Type ~ ., importance=TRUE, ntree=2000, data=fullDataSub)

  predType <- predict(fitGE)
  t1 <- table(predType, fullDataSub$Type)
  correctRate1 <- (t1[1,1]+t1[2,2]+t1[3,3])/sum(t1)
  return(correctRate1)
}

rateVector1 <- sapply(1:10, function(x) getRate1())
var(rateVector1)
```


## Decision trees using random genes
```{r}
getRate2 <- function() {
  smpl <- sample(pValueDF1$Gene, 1000)
  fullDataSample <- fullData[,c("age", "Type", "Class", smpl)] 
  fullDataSample$Type <- as.factor(as.character(fullDataSample$Type))


  fitGE2 <- randomForest(fullDataSample$Type ~ ., importance=TRUE, ntree=10000, data=fullDataSample)
  predType2 <- predict(fitGE2)
  t2 <- table(predType2, fullDataSample$Type)
  correctRate2 <- (t2[1,1]+t2[2,2]+t2[3,3])/sum(t2)
  return(correctRate2)
}

rateVector2 <- sapply(1:10, function(x) getRate2())
var(rateVector2)
```


## Decision trees without gender
```{r}
getRate3 <- function() {
  smpl1 <- sample(pValueDF1$Gene, 1000)
  fullDataSample1 <- fullData[,c("age", "Type", smpl1)] 
  fullDataSample1$Type <- as.factor(as.character(fullDataSample1$Type))

  fitGE3 <- randomForest(fullDataSample1$Type ~ ., importance=TRUE, ntree=10000, data=fullDataSample1)
  predType3 <- predict(fitGE3)
  t3 <- table(predType3, fullDataSample1$Type)
  correctRate3 <- (t3[1,1]+t3[2,2]+t3[3,3])/sum(t3)
  return(correctRate3)
}

rateVector3 <- sapply(1:10, function(x) getRate3())
var(rateVector3)
```


## Decision Trees using a random sample and save each rate in a vector
```{r}
getRate <- function() {
  samp <- sample(pValueDF1$Gene, 500)
  fullDataSamp <- fullData[,c("age", "Type", samp)] 
  fullDataSamp$Type <- as.factor(as.character(fullDataSamp$Type))

  fitGE <- randomForest(fullDataSamp$Type ~ ., importance=TRUE, ntree=500, data=fullDataSamp)
  predType <- predict(fitGE)
  t <- table(predType, fullDataSamp$Type)
  correctRate <- (t[1,1]+t[2,2]+t[3,3])/sum(t)
  return(correctRate)
}

rateVector <- sapply(1:10, function(x) getRate())
var(rateVector)
```


## Only predicting between control and asymptomatic
```{r}
getRate <- function() {
  pValueDF1$PValue <- pValueDF1$PValue[order(pValueDF1$PValue)]
  sbst <- unique(pValueDF1$Gene[1:800])
  combDatasbst <- rbind(asymptomaticData, controlsData)
  fullDatasbst <- inner_join(ageData, combDatasbst)
  fullDataSamp <- fullDatasbst[,c("age", "Class", "Type", sbst)] 
  fullDataSamp$Type <- as.factor(as.character(fullDataSamp$Type))

  fitGE <- randomForest(fullDataSamp$Type ~ ., importance=TRUE, ntree=1000, data=fullDataSamp)
  predType <- predict(fitGE)
  t <- table(predType, fullDataSamp$Type)
  correctRate <- (t[1,1]+t[2,2])/sum(t)
  return(correctRate)
}
getRate()
```


