---
title: "Colorectal Cancer Random Forest"
output: html_document
---

## Load Libraries and set working directory
```{r, echo=TRUE}
library(readr)
library(dplyr)
library(ggplot2)
library(randomForest) 

setwd("~/Dropbox/Gene_Signatures_Code")
```


## Read in data from directory & merge with meta data
```{r, echo=TRUE}
getGeneData <- function(data) {
  geneExpressionData <- read_tsv(data)
  colnames(geneExpressionData)[1] <- "SampleID"
  return(geneExpressionData)
}

metaData <- read_csv("Metadata/Ensembl87_Biomart_GRCh38.p7.csv")
colnames(metaData)[1] <- "Gene"
colnames(metaData)[2] <- "Chromosome"
colnames(metaData)[3] <- "GeneSymbol"

# Remove unnecessary column
metaData <- select(metaData, -4)

metaData <- filter(metaData, metaData$Chromosome %in% c(1:22, "X", "Y"), !is.na(metaData$GeneSymbol))
filteredMetaData <- filter(metaData, metaData$Chromosome %in% c(1:22))

metaData$Chromosome <- factor((metaData$Chromosome))
# Use filtered for barplot, regular for the dataframe
filteredMetaData$Chromosome <- factor(as.numeric(filteredMetaData$Chromosome)) 

```


## Merge with Class column
```{r}
getGenderData <- function(data) {
  genderData <- data.frame(read_tsv(data))
  genderData$Class <- factor(genderData[,2])

  genderData$Class <- ifelse(genderData$Class=="0", genderData$Class <- "Male", genderData$Class <- "Female")

  genderData <- genderData[,c(1,3)]
  return(genderData)
}
```


## Read in Data
```{r}
# Colorectal Cancer
geneExpressionData <- getGeneData("GSE26682_U133A.txt.gz")
genderData <- getGenderData("Class/U133A__gender.txt")
coloMergedData <- inner_join(genderData, geneExpressionData)

# Posterior Fossa Ependymoma
geneExpressionData <- getGeneData("GSE27279.txt.gz")
genderData <- getGenderData("Class/GSE27279__gender.txt")
epeMergedData <- inner_join(genderData, geneExpressionData)

# Smokers
geneExpressionData <- getGeneData("GSE37147.txt.gz")
genderData <- getGenderData("Class/GSE37147__gender.txt")
smokeMergedData <- inner_join(genderData, geneExpressionData)

# Lung Cancer
geneExpressionData <- getGeneData("GSE37745.txt.gz")
genderData <- getGenderData("Class/GSE37745__gender.txt")
lungMergedData <- inner_join(genderData, geneExpressionData)

# Stage I/II Colon Cancer
geneExpressionData <- getGeneData("GSE37892.txt.gz")
genderData <- getGenderData("Class/GSE37892__gender.txt")
colMergedData <- inner_join(genderData, geneExpressionData)

# Idiopathic Pulmonary Fibrosis
geneExpressionData <- getGeneData("GSE38958.txt.gz")
genderData <- getGenderData("Class/GSE38958__gender.txt")
fibMergedData <- inner_join(genderData, geneExpressionData)

# Intestine
geneExpressionData <- getGeneData("GSE40292.txt.gz")
geneExpressionData <- na.omit(geneExpressionData)
genderData <- getGenderData("Class/GSE40292__gender.txt")
intMergedData <- inner_join(genderData, geneExpressionData)

# Desmoid tumors
geneExpressionData <- getGeneData("GSE58697.txt.gz")
genderData <- getGenderData("Class/GSE58697__gender.txt")
desMergedData <- inner_join(genderData, geneExpressionData)
```


## Get p-values using ANOVA
```{r}
getPValue <- function(data) {
  nGenes <- ncol(data) - 2
  geneMatrix <- as.matrix(data[,3:nGenes + 2], ncol=nGenes)

  p <- rep(NA,nGenes)

  x <- data$Class

  for (i in 1:ncol(geneMatrix)) {
      y <- geneMatrix[,i] 
      mod <- lm(y~x)
      p[i] <- anova(mod)[1,5]
  }

  return(p)
}

```


## Associate p-values with their genes
```{r}
assignPValues <- function(data) {
  geneNames <- colnames(data)[3:(ncol(data))] 
  
  pv <- getPValue(data)
  pValueData <- data.frame(Gene=geneNames, pv)  # class
  pValueData$Gene <- as.character(pValueData$Gene)
  
  # Use filtered metadata because we don't want to use X and Y-linked genes
  pValueDF <- inner_join(filteredMetaData, pValueData)
  pValueDF <- arrange(pValueDF, pv)
  colnames(pValueDF)[4] <- "PValue"
  return(pValueDF)
}
```


## Call function to get pvalue dfs
```{r}
coloPVData <- assignPValues(coloMergedData)
epePVData <- assignPValues(epeMergedData)
smokePVData <- assignPValues(smokeMergedData)
lungPVData <- assignPValues(lungMergedData)
colPVData <- assignPValues(colMergedData)
fibPVData <- assignPValues(fibMergedData)
intPVData <- assignPValues(intMergedData)
desPVData <- assignPValues(desMergedData)

```


## Decision Tree function to get correct prediction rate
```{r}
getRate <- function(pvData, tData) {
  samp <- sample(pvData$Gene, 4000)
  dataSamp <- tData[,c("Class", samp)] 
  dataSamp$Class <- as.factor(as.character(dataSamp$Class))

  fitGE <- randomForest(dataSamp$Class ~ ., importance=TRUE, ntree=1000, data=dataSamp)
  predClass <- predict(fitGE)
  t <- table(predClass, dataSamp$Class)
  correctRate <- (t[1,1]+t[2,2])/sum(t)
  return(correctRate)
}

```


## Use getRate Function
```{r}
getRate(coloPVData, coloMergedData)
# 0.7642857
getRate(epePVData, epeMergedData)
# 0.74
getRate(smokePVData, smokeMergedData)
# 0.9385965
getRate(lungPVData, lungMergedData)
# 0.9679144
getRate(colPVData, colMergedData)
# 0.8285714
getRate(fibPVData, fibMergedData)
# 0.7565217
getRate(intPVData, intMergedData)
# 0.9825581
getRate(desPVData, desMergedData)
# 0.7704918

# rateVector <- sapply(1:10, function(x) getRate(pValueDF1, coloMergedData))
# var(rateVector)
```


## Splitting the data into training and test data
```{r}
splitdf <- function(data) {
    index <- 1:nrow(data)
    trainindex <- sample(index, length(index)*.7)
    trainset <- data[trainindex, ]
    testset <- data[-trainindex, ]
    res <- list(trainset=trainset,testset=testset)
    return(res)
}
```


## RF Using Training and Testing Data
```{r}
getRF <- function(PVData, mergedData) {
  
  ## Create training and test data
  trainandtestData <- splitdf(mergedData)
  trainData <- trainandtestData[[1]]
  testData <- trainandtestData[[2]]
  
  ## Find the important genes using training data
  samp1 <- PVData$Gene[1:trunc(nrow(PVData)/2)]
  dataSamp1 <- trainData[,c("Class", samp1)] 
  dataSamp1$Class <- as.factor(as.character(dataSamp1$Class))
  fitGE1 <- randomForest(dataSamp1$Class ~ ., importance=TRUE, ntree=1000, data=dataSamp1)
  
  samp2 <- PVData$Gene[trunc(nrow(PVData)/2):nrow(PVData)]
  dataSamp2 <- trainData[,c("Class", samp2)] 
  dataSamp2$Class <- as.factor(as.character(dataSamp2$Class))
  fitGE2 <- randomForest(dataSamp2$Class ~ ., importance=TRUE, ntree=1000, data=dataSamp2)
  
  a <- rownames(head(importance(fitGE1)))
  b <- rownames(head(importance(fitGE2)))
  impGenes <- c(a,b)

  impGenesData1 <- trainData[,c("Class", impGenes)]
  impGenesData1$Class <- as.factor(as.character(impGenesData1$Class))

  fitGE <- randomForest(impGenesData1$Class ~ ., importance=TRUE, ntree=1000, data=impGenesData1)

  predClass1 <- predict(fitGE)
  t1 <- table(predClass1, impGenesData1$Class)
  correctRate1 <- (t1[1,1]+t1[2,2])/sum(t1)
  
  ## Test Data
  predClass2 <- predict(fitGE, newdata = testData)
  
  impGenesData2 <- testData[,c("Class", impGenes)]
  impGenesData2$Class <- as.factor(as.character(impGenesData2$Class))
  
  t2 <- table(predClass2, impGenesData2$Class)
  correctRate2 <- (t2[1,1]+t2[2,2])/sum(t2)
  
  output <- c(correctRate1, correctRate2, impGenes)
  return(output)
}
```

## Call RF correct rate function
```{r}
# I started using 70-30 (training-test) for colon cancer and down, 50-50(training-test) for everything before
coloResults <- NULL
epeResults <- NULL
smokeResults <- NULL
lungResults <- NULL
colResults <- NULL
fibResults <- NULL
intResults <- NULL
desResults <- NULL

for (i in 1:5) {
  coloResults[[i]] <- getRF(coloPVData, coloMergedData)   ### 52-70%
  epeResults[[i]] <- getRF(epePVData, epeMergedData)  ### Barely more than half the time
  smokeResults[[i]] <- getRF(smokePVData, smokeMergedData)
  lungResults[[i]] <- getRF(lungPVData, lungMergedData)
  colResults[[i]] <- getRF(colPVData, colMergedData)
  fibResults[[i]] <- getRF(fibPVData, fibMergedData)
  intResults[[i]] <- getRF(intPVData, intMergedData)
  desResults[[i]] <- getRF(desPVData, desMergedData)

}
```



## OSCC (Still in progress, but putting on hold)
```{r}
geneExpressionData <- getGeneData("GSE30784.txt.gz")
geneExpressionData <- na.omit(geneExpressionData)
genderData <- getGenderData("Class/GSE30784__gender.txt")
osccMergedData <- inner_join(genderData, geneExpressionData)
# osccAgeData <- read_tsv("Class/GSE30784__age.txt")
# ageData$Age <- factor(osccAgeData$Age)
# 
# osccMergedData <- inner_join(osccAgeData, osccMergedData)

# nGenes <- ncol(osccMergedData) - 3
# geneMatrix <- as.matrix(osccMergedData[,3:nGenes + 3], ncol=nGenes)
# 
# p <- rep(NA,nGenes)
# 
# x1 <- osccMergedData$Class
# x2 <- osccMergedData$age.40.49
# 
# for (i in 1:ncol(geneMatrix)) {
#      y <- geneMatrix[,i] 
#     mod <- lm(y~x)
#     p[i] <- anova(mod)[1,5]
# }

osccPVData <- assignPValues(osccMergedData)
getRate(osccPVData, osccMergedData)
# 0.7468354
```





