---
title: "Master Code"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load libraries & set working directory

```{r, echo=TRUE}
library(readr)
library(dplyr)
library(ggplot2)
library(limma)

setwd("~/Dropbox/Gene_Signatures_Code")
```

## Read in data from directory & merge data files
```{r, echo=TRUE}
getGeneData <- function(data) {
  geneExpressionData <- read_tsv(data)
  colnames(geneExpressionData)[1] <- "SampleID"
  return(geneExpressionData)
}

metaData <- read_csv("Metadata/Ensembl87_Biomart_GRCh38.p7.csv")
colnames(metaData)[1] <- "Gene"
colnames(metaData)[2] <- "Chromosome"
colnames(metaData)[3] <- "GeneSymbol"

# Remove unnecessary column
metaData <- select(metaData, -4)

metaData <- filter(metaData, metaData$Chromosome %in% c(1:22, "X", "Y"), !is.na(metaData$GeneSymbol))
filteredMetaData <- filter(metaData, metaData$Chromosome %in% c(1:22))

metaData$Chromosome <- factor((metaData$Chromosome))
# Use filtered for barplot, regular for the dataframe
filteredMetaData$Chromosome <- factor(as.numeric(filteredMetaData$Chromosome)) 

```

## Merge Class column
```{r}
getGenderData <- function(data) {
  genderData <- data.frame(read_tsv(data))
  genderData$Class <- factor(genderData[,2])

  genderData$Class <- ifelse(genderData$Class=="0", genderData$Class <- "Male", genderData$Class <- "Female")

  genderData <- genderData[,c(1,3)]
  return(genderData)
}
```


## Create dot plot and t.test function

```{r, echo=TRUE}
dotPlot <- function(mergedData, geneOfInterest, outFilePath) {
  plotData <- select(mergedData, SampleID, Class, one_of(geneOfInterest)) #Come back to this
  colnames(plotData)[3] <- "Gene"
  plotData$Gene <- as.numeric(plotData$Gene)

  pValue <- t.test(Gene ~ Class, data = plotData)$p.value
  
  if(!is.null(outFilePath)) { 
    p <- ggplot(plotData, aes(x = Class, y = Gene)) + geom_jitter(size=.5, width=.15) +
     theme_minimal() +
     theme(axis.ticks.y = element_blank(), 
       axis.line = element_blank(),
       axis.line.y = element_line(color="gray50", size=0.2),
       panel.grid.major = element_line(color = "gray50", size = 0.2), 
       panel.grid.major.x = element_blank()) +
     ggtitle(paste(titlePrefix, geneOfInterest, paste("(", (format(pValue, trim = TRUE, digits = 2, scientific = TRUE)), ")", sep=""), sep=" ")) + 
     theme(plot.title = element_text(hjust = 0.5), x.axis.title = element_text(font = )) +
     scale_y_continuous(name = "Expression Level", seq(-6, 6, by = 0.05))
  
    print(p)

    ggsave(outFilePath, height=6, width=8, units="in")
  }
  
  return(pValue)
}
```

## Just the t.test

```{r}
getPValues <- function(mergedData) {
  nGenes <- ncol(mergedData) - 2
  geneMatrix <- as.matrix(mergedData[,3:(nGenes+2)], ncol=nGenes)
  p <- rep(NA,nGenes)
  for (i in 1:ncol(geneMatrix)) {
    y <- geneMatrix[,i]
    p[i] <- t.test(y ~ mergedData$Class)$p.value
  }
 return(p)
}
```

## Create a function to perform t.test (by calling the dotplot function) and get p-values

```{r, }
getSignificantGenes <- function(mergedData, outFilePath) {
  pValueVector <- NULL 

  for(i in 3:ncol(mergedData)) { 
    geneOfInterest <- colnames(mergedData)[i] 
    pValue <- dotPlot(mergedData, geneOfInterest, paste("Asymptomatic"), NULL) 
    
    pValueVector <- c(pValueVector, pValue)
  }
  
  geneName <- colnames(mergedData)[3:ncol(mergedData)]

  pValueDataFrame <- data.frame(Gene=geneName, PValue=pValueVector)
  pValueDataFrame <- arrange(pValueDataFrame, PValue)

  if(!is.null(outFilePath)) {
    histogram <- ggplot(pValueDataFrame, aes_(x=pValueVector)) + geom_histogram(binwidth = .05, center=.025) + ggtitle("Significant Genes") 
    print(histogram)
  
    ggsave(outFilePath, device= "pdf", height=6, width=8, units="in") 
  }
  
  return(pValueDataFrame) 
}
# Come back and figure this out so you can get a dotplot if you want but you don't need one right now.
# dotPlot(genderMergedData, genderMergedData$ENSG00000102921, paste("Figures/", "ENSG00000102921", "_DotPlot.pdf", sep = "")) 
```

## Create a function to perform t.test (by calling the getPValues function) and get p-values

```{r, }
getSignificantGenes <- function(mergedData) {
    pValue <- getPValues(mergedData) 
    geneNames <- colnames(mergedData)[3:(ncol(mergedData))] 
    o <- order(pValue)
    pValueData <- data.frame(Gene=geneNames[o],pValue=pValue[o])
    return(pValueData)
}
```

## Call Significant Genes function to get the most significant

```{r}
printGenes <- function(titlePrefix, mergedData, metaDataFrame) {

  pValueDataFrame <- getSignificantGenes(mergedData)
  pValueDataFrame$Gene <- as.character(pValueDataFrame$Gene)
  pValueDataFrame <- inner_join(metaDataFrame, pValueDataFrame)
  pValueDataFrame <- arrange(pValueDataFrame, pValue)

  return(pValueDataFrame) 
}
```


## create a function that calls print genes function and gets the proportion of significant genes for each chromosome
```{r}
createNewDF <- function(titlePrefix=NULL, mergedData, metaDataFrame) {
  
  genderData <- printGenes(titlePrefix, mergedData, metaDataFrame)

  newDataFrame <- subset(genderData, select=c("Chromosome", "Gene", "pValue"))

  for(i in 1:nrow(newDataFrame)) {
    ifelse(newDataFrame$pValue[i]<=.05, newDataFrame$pValue[i] <- "Yes", newDataFrame$pValue[i] <- "No")
  }

  df <- data.frame(newDataFrame)
  return(df)
  return(print(hist(genderData$pValue)))
}
```

# If you want to make a barplot, use this function instead of the createNewDF function
```{r}
createBarPlot <- function(titlePrefix, mergedData, metaDataFrame) {
  genderData <- printGenes(titlePrefix, mergedData, metaDataFrame)
  
  newDataFrame <- subset(genderData, select=c("Chromosome", "Gene", "pValue"))

  for(i in 1:nrow(newDataFrame)) {
    ifelse(newDataFrame$pValue[i]<=.05, newDataFrame$pValue[i] <- "Yes", newDataFrame$pValue[i] <- "No")
  }

  df <- data.frame(newDataFrame)
  
  group_by(df, Chromosome, pValue)
  summarise(df, count = n())

  propPlot <- ggplot(df, aes(x=Chromosome, fill=pValue)) + geom_bar() + ggtitle(paste(titlePrefix, "Significant Genes", by=" "))
  return(propPlot)
}

```

# After Gastric dataset. Talk to Dr. Piccolo, one of his grad students or someone who can tell us if it's paper-worthy.
# Continue on to all data sets, any patterns? See anything?
# If not, is there a different question we can ask?
# Look at other papers about gender genetic differences



## Look up prevalance of the disease in females
## Check which genes are significant. Are the same ones appearing?
## Look if they have other classifications to do ^^^
## Send the list of diseases to Dr. Tass.
## Is the lung file in the wrong format? It's not reading in any data
## Make a summary that has the most significnace
## Make a table of the actual genes that are significant

## Colorectal and Colon cancers are related
## Work on Presentation!!!
## Machine Learning. If I look at the gene signature can I predict what disease they have--especially helpful if they're asymptomatic. Early detection.
## Ask Dr. Piccolo what Haley was doing.
## First: SUMMARIZE. Did we answer any questions? What questions do we still have that we can concretely answer?

