---
title: "Using the Master Code"
output: html_document
---

## Colorectal Cancer
```{r}
geneExpressionData <- getGeneData("GSE26682_U133A.txt.gz")
genderData <- getGenderData("Class/U133A__gender.txt")
coloMergedData <- inner_join(genderData, geneExpressionData)

colorectalNewDF <- createNewDF(NULL, coloMergedData, metaData)
createBarPlot(paste("Colorectal", "Cancer", by=" "), coloMergedData, filteredMetaData)
```

## Posterior Fossa Ependymoma
```{r}
geneExpressionData <- getGeneData("GSE27279.txt.gz")
genderData <- getGenderData("Class/GSE27279__gender.txt")
epeMergedData <- inner_join(genderData, geneExpressionData)

epeNewDF <- createNewDF(epeMergedData, metaData)
# Use filtered because X and Y are not numeric
createBarPlot("Ependymoma", epeMergedData, filteredMetaData)

```

## Lung Cancer (Not working?)
```{r}
geneExpressionData <- getGeneData("GSE2109_Lung.txt.gz")
genderData <- getGenderData("Class/Lung__Gender.txt")
lungMergedData <- inner_join(genderData, geneExpressionData)

lungNewDF <- createNewDF(paste("Lung", "Cancer", by=" "), lungMergedData, metaData)
```

## Smokers
```{r}
geneExpressionData <- getGeneData("GSE37147.txt.gz")
genderData <- getGenderData("Class/GSE37147__gender.txt")
smokeMergedData <- inner_join(genderData, geneExpressionData)

smokeNewDF <- createNewDF("Smoker", smokeMergedData, metaData)
createBarPlot("Smoker", smokeMergedData, filteredMetaData)

```

## Lung Cancer (Not working?)
```{r}
geneExpressionData <- getGeneData("GSE30219.txt.gz")
genderData <- getGenderData("Class/GSE30219__gender.txt")
lungMergedData <- inner_join(genderData, geneExpressionData)

lungNewDF <- createNewDF("Lung", lungMergedData, metaData)
```

## Oral Squamous Cell Carcinoma (OSCC)
```{r}
geneExpressionData <- getGeneData("GSE30784.txt.gz")
genderData <- getGenderData("Class/GSE30784__gender.txt")
osccMergedData <- inner_join(genderData, geneExpressionData)
# osccAgeData <- read_tsv("Class/GSE30784__age.txt")
# ageData$Age <- factor(osccAgeData$Age)
# 
# osccMergedData <- inner_join(osccAgeData, osccMergedData)
# createBarPlot("OSCC", osccMergedData[,-c(1:3)], filteredMetaData)

osccNewDF <- createNewDF("OSCC", osccMergedData, metaData)
createBarPlot("OSCC", osccMergedData, filteredMetaData)

## Check this histogram bc a lot are significant
osccGenderGenes <- printGenes("OSCC", osccMergedData, metaData)
hist(osccGenderGenes$pValue)

osccSigGenes <- osccNewDF[which(osccNewDF$pValue=="Yes"),]
```
# I want to compare females with oscc to females with other things

## Lung Cancer
```{r}
geneExpressionData <- getGeneData("GSE37745.txt.gz")
genderData <- getGenderData("Class/GSE37745__gender.txt")
lungMergedData <- inner_join(genderData, geneExpressionData)

lungNewDF <- createNewDF("Lung", lungMergedData, metaData)
createBarPlot("Lung", lungMergedData, filteredMetaData)
# This is also 20-30% significant

## Check this histogram bc a lot are significant
lungGenderGenes <- printGenes("Lung", lungMergedData, metaData)
hist(lungGenderGenes$pValue)

lungSigGenes <- lungNewDF[which(lungNewDF$pValue=="Yes"),]
```


## Stage II and III Colon Cancer
```{r}
geneExpressionData <- getGeneData("GSE37892.txt.gz")
genderData <- getGenderData("Class/GSE37892__gender.txt")
colMergedData <- inner_join(genderData, geneExpressionData)

colNewDF <- createNewDF("Colon", colMergedData, metaData)
createBarPlot("Colon", colMergedData, filteredMetaData)
# About 10% significant

# colGenderGenes <- printGenes("Colon", colMergedData, metaData)
# hist(colGenderGenes$pValue)
```

## Idiopathic Pulmonary Fibrosis
```{r}
geneExpressionData <- getGeneData("GSE38958.txt.gz")
genderData <- getGenderData("Class/GSE38958__gender.txt")
fibMergedData <- inner_join(genderData, geneExpressionData)

fibNewDF <- createNewDF("Fibrosis", fibMergedData, metaData)
createBarPlot("Fibrosis", fibMergedData, filteredMetaData)

## Check this histogram bc a lot are significant
fibGenderGenes <- printGenes("Fibrosis", fibMergedData, metaData)
hist(fibGenderGenes$pValue)

fibSigGenes <- fibNewDF[which(fibNewDF$pValue=="Yes"),]

```

## Colon Cancer (Not working)
```{r}
geneExpressionData <- getGeneData("GSE39582.txt.gz")
genderData <- getGenderData("Class/GSE39582__gender.txt")
col2MergedData <- inner_join(genderData, geneExpressionData)

col2NewDF <- createNewDF("Colon", col2MergedData, metaData)
createBarPlot("Colon", col2MergedData, filteredMetaData)

```

## Intestine
```{r}
 
intNewDF <- createNewDF("Intestine", intMergedData, metaData)
createBarPlot("Intestine", intMergedData, filteredMetaData)
```

## Desmoid Tumors
```{r}
geneExpressionData <- getGeneData("GSE58697.txt.gz")
genderData <- getGenderData("Class/GSE58697__gender.txt")
desMergedData <- inner_join(genderData, geneExpressionData)

desNewDF <- createNewDF("Desmoid", desMergedData, metaData)
createBarPlot("Desmoid", desMergedData, filteredMetaData)

desGenderGenes <- printGenes("Desmoid", desMergedData, metaData)
hist(desGenderGenes$pValue)
```

## ATTR (Symptomatic)
```{r}
geneExpressionData <- getGeneData("GSE67784.txt.gz")
genderData <- getGenderData("Class/GSE67784__gender.txt")
attrMergedData <- inner_join(genderData, geneExpressionData)

attrNewDF <- createNewDF("ATTR", attrMergedData, metaData)
createBarPlot("ATTR", attrMergedData, filteredMetaData)

attrGenderGenes <- printGenes("ATTR", attrMergedData, metaData)
hist(attrGenderGenes$pValue)

attrSigGenes <- attrNewDF[which(attrNewDF$pValue=="Yes"),]

```


## Comparing significant genes for oscc and lung cancer
```{r}
allSigGenes <- inner_join(lungSigGenes,osccSigGenes)
allSigGenes <- inner_join(allSigGenes, fibSigGenes)
allSigGenes <- inner_join(allSigGenes, attrSigGenes)
```
