---
title: "Amyloidosis Gene Expression"
output: html_document
editor_options: 
  chunk_output_type: console
---


## Load libraries & set working directory

```{r, echo=TRUE}
library(readr)
library(dplyr)
library(ggplot2)

setwd("~/Dropbox/Gene_Signatures_Code")
```

## Read in gene expression data from directory & merge with meta data 

```{r, echo=TRUE}
geneExpressionData <- read_tsv("GSE67784.txt.gz")
colnames(geneExpressionData)[1] <- "SampleID"

metaData <- read_csv("Metadata/Ensembl87_Biomart_GRCh38.p7.csv")
colnames(metaData)[1] <- "Gene"
colnames(metaData)[2] <- "Chromosome"
colnames(metaData)[3] <- "GeneSymbol"

# Remove unnecessary column
metaData <- select(metaData, -4)

metaData <- filter(metaData, metaData$Chromosome %in% c(1:22, "X", "Y"), !is.na(metaData$GeneSymbol))
# filteredMetaData <- filter(metaData, metaData$Chromosome %in% c(1:22))

metaData$Chromosome <- factor(metaData$Chromosome)
# filteredMetaData$Chromosome <- factor(filteredMetaData$Chromosome)
```

## Write a function to create separate datasets for Type (Asymptomatic, Symptomatic...)

```{r,}
getClassData <- function(geneExpressionData, classFile) {
  classData <- read_tsv(classFile)
  classData$Class <- factor(classData$Class)

  mergedData <- inner_join(classData, geneExpressionData)

  return(mergedData)
}
```

## Create separate datasets

```{r}
asymptomaticMergedData <- getClassData(geneExpressionData, "Class/Asymptomatic__gender.txt")
symptomaticMergedData <- getClassData(geneExpressionData, "Class/Symptomatic__gender.txt")
treatedMergedData <- getClassData(geneExpressionData, "Class/Treated__gender.txt")
onlycontrolsMergedData <- getClassData(geneExpressionData, "Class/Only_Controls__gender.txt")
```

## Create t.test function

```{r, echo=TRUE}
getPValues <- function(mergedData) {
  nGenes <- ncol(mergedData) - 2
  geneMatrix <- as.matrix(mergedData[,3:(nGenes+2)], ncol=nGenes)
  p <- rep(NA,nGenes)
  for (i in 1:ncol(geneMatrix)) {
    y <- geneMatrix[,i]
    p[i] <- t.test(y ~ mergedData$Class)$p.value
  }
 return(p)
}
```

## Create a function to perform t.test (by calling the getPValues function) and get p-values

```{r, }
getSignificantGenes <- function(mergedData) {
    pValue <- getPValues(mergedData) 
    geneNames <- colnames(mergedData)[3:(ncol(mergedData))] 
    o <- order(pValue)
    pValueData <- data.frame(Gene=geneNames[o],pValue=pValue[o])
    return(pValueData)
}

```

## Call Significant Genes function to get the the genes with the most significant p-values 

```{r}
printGenes <- function(titlePrefix, mergedData, metaDataFrame) {

  title <- paste(titlePrefix, "_Results_topTen", sep = "")

  pValueDataFrame <- getSignificantGenes(mergedData)
  pValueDataFrame$Gene <- as.character(pValueDataFrame$Gene)
  pValueDataFrame <- inner_join(metaDataFrame, pValueDataFrame)
  pValueDataFrame <- arrange(pValueDataFrame, pValue)
  return(pValueDataFrame) 
}
```

## Call print genes function

```{r}
asymptomaticGenes <- printGenes("Asymptomatic", asymptomaticMergedData, metaData) 
symptomaticGenes <- printGenes("Symptomatic", symptomaticMergedData, metaData)
treatedGenes <- printGenes("Treated", treatedMergedData, metaData)
onlycontrolsGenes <- printGenes("Only_Controls", onlycontrolsMergedData, metaData)

```

## Create a function that finds the proportion of significant genes for each chromosome

```{r}
createNewDF <- function(genesData) {
  newDataFrame <- subset(genesData, select=c("Chromosome", "Gene", "pValue"))

  for(i in 1:nrow(newDataFrame)) {
    ifelse(newDataFrame$pValue[i]<=.05, newDataFrame$pValue[i] <- "Yes", newDataFrame$pValue[i] <- "No")
  }

  df <- data.frame(newDataFrame)
  return(df)
}
```

## Create a function which makes a bar plot of the proportion

```{r}
createBarPlot <- function(genesDF, prefix) {
 
  group_by(genesDF, Chromosome, pValue)
  summarise(genesDF, count = n())

   propPlot <- ggplot(genesDF, aes(x=Chromosome, fill=pValue)) + geom_bar() + ggtitle(paste(prefix, "Significant Genes", by=" "))
   return(propPlot)
  
  return(genesDF)
}

```

## Call the function to find the proportion

```{r}
asymptomaticNewDF <- createNewDF(asymptomaticGenes)
symptomaticNewDF <- createNewDF(symptomaticGenes)
treatedNewDF <- createNewDF(treatedGenes)
onlycontrolsNewDF <- createNewDF(onlycontrolsGenes)
```

## Call the function to print the proportion bar plot

```{r}
createBarPlot(asymptomaticNewDF, "Asymptomatic")
createBarPlot(symptomaticNewDF, "Symptomatic")
createBarPlot(treatedNewDF, "Treated")
createBarPlot(onlycontrolsNewDF, "Controls")
```

## Do some extra calculations to find the genes which are significant between all amyloidosis patients
```{r}
impGenes <- inner_join(asymptomaticNewDF, symptomaticNewDF)
impGenes2 <- inner_join(impGenes, treatedNewDF)

# Frequency of significant genes between all amyloidosis patients
freqImpGenes <- impGenes2 %>% 
  count(Chromosome, pValue) %>%
  group_by(Chromosome) %>%
  mutate(freq = n / sum(n))

impGenes3 <- impGenes2[which(impGenes2$pValue=="Yes"),]

# Find significantly different genes for those without amyloidosis 
ocImpGenes <- onlycontrolsNewDF[which(onlycontrolsNewDF$pValue=="Yes"),]

# Significantly different genes for males and females with and without
impGenes4 <- inner_join(impGenes2, onlycontrolsNewDF)
allImpGenes <- impGenes4[which(impGenes4$pValue=="Yes"),]

# Compare these as we continue to the other datasets
nrow(impGenes3)
nrow(ocImpGenes)
nrow(allImpGenes)
```

## Plot histogram of p-values

```{r}
hist(asymptomaticGenes$pValue, main=paste("Asymptomatic", "P-Values"), xlab = "P-Values")
hist(symptomaticGenes$pValue, main=paste("Symptomatic", "P-Values"), xlab = "P-Values")
hist(treatedGenes$pValue, main=paste("Treated", "P-Values"), xlab = "P-Values")
hist(onlycontrolsGenes$pValue, main=paste("Controls", "P-Values"), xlab = "P-Values")
```


